Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-03-23 Report/kkt_mse_approach.tex
updated

To avoid the back-haul exchanges between the \acp{BS}, as an alternative approach, users perform all processing required and \acp{BS} will update the precoders based on the feedback information from the users. Upon receiving the transmit precoders from \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\lambda_{l,k,n}} and \me{\alpha_{l,k,n}}. The updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilot symbols with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. On receiving the precoded uplink pilots, each \ac{BS} use the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} in \eqref{kkt-mse-4.3} to update the transmit precoders, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Finally, Algorithm \ref{algo-4} outlines the distributed precoder design using the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
 \ForEach{\ac{BS} \me{b \in \mc{B}}}{
 initialize \me{i = 0} \;
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform downlink transmission \;
	find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
	evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
	using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
	\me{i = i + 1} \;
 }
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

>>>>

To avoid the back-haul exchanges between the \acp{BS}, as an alternative approach, users perform all processing required and \acp{BS} will update the precoders based on the feedback information from the users. Upon receiving the transmit precoders from \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\lambda_{l,k,n}} and \me{\alpha_{l,k,n}}. The updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilot symbols with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. On receiving the precoded uplink pilots, each \ac{BS} use the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} in \eqref{kkt-mse-4.3} to update the transmit precoders, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Finally, Algorithm \ref{algo-4} outlines the distributed precoder design using the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
 \ForEach{\ac{BS} \me{b \in \mc{B}}}{
 initialize \me{i = 0} \;
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform precoded downlink pilot transmission \;
	find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
	evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
	using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
	\me{i = i + 1} \;
 }
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

### 2015-03-23 Report/simulation_distributed.tex
revised
updated nam comments review should be modified
updated

The distributed algorithms are compared using the total number of backlogged packets after each \ac{SCA} update points. \review{Fig. \ref{fig-d} compares the performance of the algorithms for the configuration \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace} with \me{N_T = 4} transmit antennas at the \acp{BS} with the \ac{PL} varies uniformly between \me{[0,-6]} dB}. Each \ac{BS} serves \me{|\mc{U}_b| = 4} users in a coordinated manner to reduce the number of backlogged packets at each \ac{BS}. The number of queued packets is given by \me{Q_k = [5,7,9,11,8,12,5,4]} bits. As discussed in Section \ref{sec-4}, the performance and the convergence speed of the distributed algorithms are susceptible to the step size used in the subgradient update. Due to the fixed interference levels in the primal approach, it may lead to infeasible solutions if the initial or any intermediate update is not feasible.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{fig-3-2}
	\caption{Convergence of the centralized and the distributed algorithms for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace} \review{using \me{\ell_1} norm for \ac{JSFRA} schemes}}
	\label{fig-d}
\end{figure}

Fig. \ref{fig-d} plots the performance of the primal and the \ac{ADMM} solutions for the \ac{JSFRA} scheme using the \ac{SCA} and by \ac{MSE} relaxation at each \ac{SCA} point. In between the \ac{SCA} updates, the primal or the \ac{ADMM} scheme is performed for \me{J_{\max} = 20} iterations to exchange the respective coupling variables. In Fig. \ref{fig-d}, the total number of backlogged packets at each \ac{SCA} points are plotted without the inner loop iterations of \me{J_{\max}} times for the primal or the dual variables convergence. It can be seen from Fig. \ref{fig-d} that the distributed algorithms approach the centralized performance by exchanging minimal information between the coordinating \acp{BS}.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{fig-9-3}
	\caption{Impact of varying \me{q} in the total number of backlogged packets after each \ac{SCA} update for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
	\label{fig-d-3.1}
\end{figure}

\review{Fig. \ref{fig-d-3.1} compares the performance of the centralized and the \ac{KKT} algorithm in Section \ref{sec-4.3} for different exponents with \me{Q_k = [9,16,14,16,9,13,11,12]} bits and the \ac{PL} varies uniformly between \me{[0,-3]} dB}. The \me{\ell_1} norm \ac{JSFRA} scheme provides better performance over other schemes due to the greedy objective. The \ac{KKT} approach for \me{\ell_1} norm is not defined due to the non-differentiability of the objective as discussed in the Section \ref{sec-4.3}. If used for \me{\ell_1} norm, the problem of over-allocation will not affect the dual variables \me{\sigma_{l,k,n}} and \me{\alpha_{l,k,n}} since the queue deviation is raised to the power zero in \eqref{kkt-mse-4.2}, which will always be equal to one. A heuristic method based on subdifferential calculus in \cite{bertsekas1999nonlinear} is proposed in Fig. \ref{fig-d-3.1} by assigning zero for \me{\sigma_{l,k,n}} when the queue deviation is negative, \textit{i.e}, \me{Q_k - t_k < 0}. It is required to address the problem of over-allocation in the \me{\ell_1} norm for dropping the absolute value operator from the objective function. It can be seen that the heuristic method oscillates near the stationary point with the deviation determined by the factor \me{\rho} used in \eqref{kkt-mse-4.1}.
\begin{figure*}
	\centering
	\subfloat[][Average backlogged packets in the system after \me{100} transmission instants]{
		\includegraphics[width=0.48\textwidth]{average_queue_over_time}
		\label{fig-review}
	}
	\hfill
	\subfloat[][Total backlogged packets at each transmission slot for \me{A_k = 5} bits]{
		\includegraphics[width=0.48\textwidth]{instant_queue_over_time}
		\label{fig-review-time}
	}
	\caption{Time analysis of the Queue dynamics for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace}}
	\label{fig-time-analysis}
\end{figure*}

The objective values are mentioned in the legend for all the schemes and the objective of the \me{\ell_2} norm is not the same as that of the \me{\ell_1} norm used for plotting. For simulations, we update all variables in \eqref{kkt-mse-4} at once at each iteration, \textit{i.e}, \me{J_{\max} = 1}, which is well justified for the practical implementations due to the signaling overheads. The \me{\ell_2} norm for the \ac{JSFRA} and the \ac{KKT} approach achieves nearly the same value of \me{6.62} with different \me{\chi}, due to the limited number of iterations for the dual variable convergence between each \ac{SCA} update. Fig. \ref{fig-d-3.1} also shows the effect of dropping the squared rate variable from the objective in the \ac{Q-WSRME} scheme compared to the \me{\ell_2} norm which includes it. By dropping it, the \ac{Q-WSRME} scheme minimizes the number of queued packets in a prioritized manner based on the respective queues. On contrary, the \me{\ell_2} norm allocate rates to the users with the higher number of queued packets before addressing the users with the smaller number of queued packets.

>>>>

The distributed algorithms are compared using the total number of backlogged packets after each \ac{SCA} update. \review{Fig. \ref{fig-d} compares the performances of the algorithms with the \ac{PL} varies uniformly between \me{[0,-6]} dB}. Each \ac{BS} serves \me{|\mc{U}_b| = 4} users in a coordinated manner to reduce the number of backlogged packets, whith the queued packets for each user is given by \me{Q_k = [5,7,9,11,8,12,5,4]} bits. As discussed in Section \ref{sec-4}, the performance and the convergence speed of the distributed algorithms are susceptible to the step size used in the subgradient update. Due to the fixed interference levels in the primal approach, it may lead to infeasible solutions if the initial or any intermediate update is not feasible.

Fig. \ref{fig-d} compares the primal and the \ac{ADMM} solutions for the \ac{JSFRA} scheme using the \ac{SCA} and by \ac{MSE} relaxation using the number of backlogged packets. In between the \ac{SCA} updates, the primal or the \ac{ADMM} scheme is performed for \me{J_{\max} = 20} iterations to exchange the respective coupling variables. The number of backlogged packets only at the \ac{SCA} points are marked in the figure. The performance of the distributed approaches are similar to the centralized schemes if the primal and dual updates are performed until convergence.
%In Fig. \ref{fig-d}, the total number of backlogged packets at each \ac{SCA} points are plotted without the inner loop iterations of \me{J_{\max}} times for the primal or the dual variables convergence. %It can be seen from Fig. \ref{fig-d} that the distributed algorithms approach the centralized scheme by exchanging minimal information between the coordinating \acp{BS}.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{fig-9-3}
	\caption{Impact of varying \me{q} in the total number of backlogged packets after each \ac{SCA} update for a system \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 5,2,8,4,1 \rbrace}}
	\label{fig-d-3.1}
\end{figure}

\review{Fig. \ref{fig-d-3.1} compares the performances of the centralized and the \ac{KKT} algorithm in Section \ref{sec-4.3} for different exponents with \me{Q_k = [9,16,14,16,9,13,11,12]} bits and the \ac{PL} varies uniformly between \me{[0,-3]} dB}. The \me{\ell_1} norm \ac{JSFRA} scheme performs better over other schemes due to the greedy objective. The \ac{KKT} approach for \me{\ell_1} norm is not defined due to the non-differentiability of the objective as discussed in the Section \ref{sec-4.3}. If used for \me{\ell_1} norm, the problem of over-allocation will not affect the dual variables \me{\sigma_{l,k,n}} and \me{\alpha_{l,k,n}} since the queue deviation is raised to the power zero in \eqref{kkt-mse-4.2}. A heuristic method based on subdifferential calculus in \cite{bertsekas1999nonlinear} is proposed in Fig. \ref{fig-d-3.1} by assigning zero for \me{\sigma_{l,k,n}} when \me{Q_k - t_k < 0}. It addresses the over-allocation in the \me{\ell_1} norm for dropping the absolute value operator from the objective. It can be seen that the heuristic method oscillates near the converging point with the deviation determined by the factor \me{\rho} used in \eqref{kkt-mse-4.1}. The objective values are mentioned in the legend for all the schemes and the objective of the \me{\ell_2} norm is not the same as that of the \me{\ell_1} norm used for plotting.
\begin{figure*}
	\centering
	\subfloat[][Average backlogged packets in the system after \me{250} transmission instants]{
		\includegraphics[width=0.48\textwidth]{average_queue_over_time-2}
		\label{fig-review}
	}
	\hfill
	\subfloat[][Total backlogged packets at each transmission slot for \me{A_k = 6} bits]{
		\includegraphics[width=0.48\textwidth]{instant_queue_over_time-2}
		\label{fig-review-time}
	}
	\caption{Time analysis of the Queue dynamics for a system \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 4,2,12,4,1 \rbrace}}
	\label{fig-time-analysis}
\end{figure*}
%The \me{\ell_2} norm for the \ac{JSFRA} and the \ac{KKT} approach achieves nearly the same value of \me{6.62} with different \me{\chi}, due to the limited number of iterations for the dual variable convergence between each \ac{SCA} update. Fig. \ref{fig-d-3.1} also shows the effect of dropping the squared rate variable from the objective in the \ac{Q-WSRME} scheme compared to the \me{\ell_2} norm which includes it. By dropping it, the \ac{Q-WSRME} scheme minimizes the number of queued packets in a prioritized manner based on the respective queues. On contrary, the \me{\ell_2} norm allocate rates to the users with the higher number of queued packets before addressing the users with the smaller number of queued packets.
